### 2023-08-11 thesis/sections/4-exp-setup.tex
restructure thesis

\subsection{Creation of the dataset}
The basis for my datasets is the CLEVR dataset \citep{Johnson2016}.
This dataset includes 3D-generated images depicting scenes with different kinds of objects.
Each of these objects has different combinations attributes, such as \emph{shape}, \emph{color}, \emph{size} and \emph{material}.
The possible values of these attributes are listed in Table \ref{tab:clevr-attributes}.
Three to ten objects are randomly placed into the scene and assigned with random attributes.
To enhance realism and reduce ambiguity, objects do not intersect, have a certain distance from each other, and are at least partially visible.
Figure \ref{fig:clevr-example} shows an example of a generated image in the CLEVR dataset.

\begin{table}[h]
    \centering
    \begin{tabular}{cccc}
        \toprule
        \textbf{ shape } & \textbf{ color } & \textbf{ size } & \textbf{ material } \\
        cube             & gray             & small           & rubber              \\
        sphere           & red              & large           & metal               \\
        cylinder         & blue                                                     \\
                         & green                                                    \\
                         & brown                                                    \\
                         & purple                                                   \\
                         & cyan                                                     \\
                         & yellow                                                   \\
        \bottomrule
    \end{tabular}
    \caption{Attributes of objects in the CLEVR dataset}
    \label{tab:clevr-attributes}
\end{table}

\begin{figure}[h]
    \centering
    \includegraphics[width=.8\linewidth]{figures/CLEVR_example.png}
    \caption{Example of a generated image in the CLEVR dataset}
    \label{fig:clevr-example}
\end{figure}

Furthermore, the dataset contains information about each scene.
This includes all selected attributes for each object as well as the exact position of the centers of all the objects, both 3D-coordinates in the 3D scene and 2D-coordinates in the final rendered image.
In addition, simple spatial relations (in front of, behind, left, right) between the objects are calculated and stored.
These are simply based on the 3D-coordinates of the objects in relation to the position of the camera.

\cmtDK{ambiguity}
\cmtDK{referring expressions}

@@

\subsection{Description of experiments}
Language games are very complex setups for machine learning models.
The models need to solve multiple tasks at the same time to solve the overall problem.
For instance, in a simple setup of a game two agents are involved.
The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own \cmtDK{?}{space}.
In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
And finally it needs to learn how to identify the target object with this information.

@@

\subsubsection{Feature extractors}
To extract features from the images, I make use of different deep neural network architectures that are developed for this task.
First, I use the \cmtDK{why ResNet/VGG?} VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
After the convolutional layers, the data is passed first through an average pooling layer which outputs 512x7x7 dimensions.
Next follow three linear layers with \emph{ReLU} non-linearities in between.
After flattening the input, these classification layers output 4069, 4069 and 1000 dimensions respectively.

Secondly, I include the ResNet-101 \citep{He2016} that tries to overcome problems of very deep networks, using \cmtDK{needs to be explained deeper?}{residual blocks}.
There are four blocks that output 256x56x56, 512x28x28, 1024x14x14 and 2048x1x1 dimensions.
A following average pooling layer outputs 2048x1x1 dimensions as well.
The final linear layer reduces the flattened data to 1000 dimensions, corresponding to the ImageNet classes.

Both architectures are available pretrained on an image classification task on the ImageNet dataset.
Since the task in this research is very different from a classification, multiple different adaptions of these architectures are compared. \cmtDK{deeper discussion about different layers of these networks}
Table \ref{tab:feature-extractor-archs} lists the different adaptions for both VGG19 and ResNet-101 that will be used in this research.
In the later chapters, it will be referred to these adaptions using the name in the table.

\begin{table}[h]
    \centering
    \begin{tabular}{rlc}
        \toprule
                            & \textbf{description}                                            & \textbf{ output dimensions } \\\midrule
        \textbf{VGG-0}      & contains only the convolutional layers                          & 512x7x7                      \\
        \textbf{VGG-avg}    & contains an additional average pooling layer                    & 512x7x7                      \\
        \textbf{VGG-cls1}   & \makecell[cl]{contains an additional one classification layer,                                 \\ including its non-linearity} & 4069                         \\
        \textbf{VGG-cls2}   & \makecell[cl]{contains another additional classification layer,                                \\ including its non-linearity} & 4069                         \\
        \textbf{VGG-cls3}   & the original VGG19 architecture                                 & 1000                         \\\midrule
        \textbf{ResNet-1}   & contains one residual block                                     & 256x56x56                    \\
        \textbf{ResNet-2}   & contains two residual blocks                                    & 512x28x28                    \\
        \textbf{ResNet-3}   & contains three residual blocks                                  & 1024x14x14                   \\
        \textbf{ResNet-4}   & contains four residual blocks                                   & 2048x1x1                     \\
        \textbf{ResNet-avg} & contains an additional average pooling layer                    & 2048x1x1                     \\
        \textbf{ResNet-cls} & the original ResNet-101 architecture                            & 1000                         \\
        \bottomrule
    \end{tabular}
    \caption{Different adaptions of VGG19 and ResNet-101 used in this research}
    \label{tab:feature-extractor-archs}
\end{table}

Furthermore, I experimented with both the pretrained models as well as with the architectures trained from scratch with a random initialization for the weights.
This reason for this was to test if the success of an experiment was actually making use of the pretrained knowledge incorporated in the models.
If that was not the case, the agents were likely not using image features, but instead relying on some other underlying patterns to solve the task.
Basically, this approach works as an indicator to determine the actual success of the agents aside from measures as the accuracy or precision.

\subsubsection{Pre-experiments without language games}
The final objective of this thesis is to find out, how agents can communicate about relations of objects spatially as well as based on their attributes.
Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge.
Here, I structured the experiments into three levels.
In the first level, the model are trained to learn the position of objects in the image and attend to specific regions of the image.
In the second level, the models are trained to differentiate objects in the scene from each other.
The last level adds language to the experiment, more specifically the models learn to caption and describe objects in the image.
These combined experiments should lay the basis for how to build up the agents in the language games.

\paragraph{Level 1 - Coordinate predictor:}

@@

\paragraph{Level 2 - Bounding box classifier:}
\cmtDK{better to call it discriminators?}

One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
Feature extractors like VGG or ResNet are trained on extracting visual information.
In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
To address this challenge, I restructured the task in this experiment.
Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
To do this, fixed-size bounding boxes are extracted around each of the object in the image.
As for the masking, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
The bounding boxes of all objects are padded to a number of 10 bounding boxes, the maximum number of objects in an image with a matrix of zeros.
These bounding boxes are shuffled.

The model passes each of the bounding boxes through a feature extractor.
The resulting vectors are concatenated, which basically is a concatenation of the extracted features of each object.
This long vector is passed through a linear layer with 4096 with a subsequent \emph{ReLU} function.
Afterwards, a linear layer reduces the dimensions to 10, corresponding to the 10 bounding boxes in the input.
With a final \emph{softmax} function, the model then points to one of the bounding boxes as the target bounding box.
The loss is calculated using cross entropy.

@@

\paragraph{Level 3 - Caption generator:}

@@

\subsubsection{Language games}
The experiments that are executed using language games have a similar structure as the pre-experiments, since those lied the basis for the language games.

@@

\paragraph{Discrimination games:}
In a discrimination game, the agents are presented with two or more images, one of these being the target image.
The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
The receiver then needs to decide based on the message, which of the images is the target image.

@@

\paragraph{Caption generators:}
In a next step, it is tested if the agents can learn to extract features of the objects together.
For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
The receiver only sees the image without additional information.
\cmtDK{human bias vs emergent language}

@@

>>>>

\subsection{Creation of the dataset}
\cmtDK{ambiguity}
\cmtDK{referring expressions}

@@

\subsection{Generating and understanding referring expressions}
Language games are very complex setups for machine learning models.
The models need to solve multiple tasks at the same time to solve the overall problem.
For instance, in a simple setup of a game two agents are involved.
The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own \cmtDK{?}{space}.
In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
And finally it needs to learn how to identify the target object with this information.

@@

The final objective of this thesis is to find out, how agents can communicate about relations of objects spatially as well as based on their attributes.
Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge.
Here, I structured the experiments into three levels.
In the first level, the model are trained to learn the position of objects in the image and attend to specific regions of the image.
In the second level, the models are trained to differentiate objects in the scene from each other.
The last level adds language to the experiment, more specifically the models learn to caption and describe objects in the image.
These combined experiments should lay the basis for how to build up the agents in the language games.

\subsubsection{Coordinate predictor}

@@

\paragraph*{Results:}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
    \label{fig:coordinate-predictor_loss}
\end{figure}

Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ significantly with other feature extractors.
As can be seen, the success between the different datasets differ significantly.
The more objects are present in an image, the worse the model performs.
The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
This prediction even though not perfectly on the center point is always on the object.
Opposed to that, using the Dale-2 dataset with two objects, the mean distance is from the first epoch between 37 and 38 pixels.
With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.

An interesting observation is the difference of the mean distances between training and testing data.
The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
This points to the fact that the model is not generalizing the task and learning abstract patterns that can be applied to unseen data, but instead learning the training data by heart.
That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
Applying a higher dropout didn't have an impact on the results.

This behavior is indeed not very surprising.
First, these results show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
Secondly, more than one objects present in an image confuse the model, and it is not able to consistently point to one of them.
This can have multiple reasons, for instance that models lacks the ability to separate objects in the extracted features.
But even given, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
The guess is then more or less random.
This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
This task on the other hand seems very difficult to learn.
In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.

When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
One factor that now has a much higher impact is the feature extractor that is used.
Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
First, a big difference can be seen between the datasets.
The model only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
In most cases, it stays above 50 pixels.
Using the Dale-2 dataset, the behavior is a little different.
All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
This effect also applies when using less residual blocks.
Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.

\begin{table}[h]
    \centering
    \begin{tabular}{rcccc}
        \toprule
                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
                            & train                               & test                                & train          & test           \\\midrule
        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
        \bottomrule
    \end{tabular}
    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
    \label{tab:feature-extractor-mean-distances}
\end{table}

Secondly, the training loss now looks also different.
In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
The only exception is ResNet-3 as a feature extractor.
In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
Higher layers, with more specific encoded information need to be used for this research.
The experiments in the following sections are set up using these higher layers.

Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
This hypothesis doesn't hold.
All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
The model is still not able to understand, which object is the target object.
For that reason, a better representation of the target object is necessary.

In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.

\begin{figure}[h]
    \centering
    \subfigure['Dale-2', train split]{
        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
        \label{fig:visualizations_dale-2_train}
    }
    \subfigure['Dale-2', test split]{
        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
        \label{fig:visualizations_dale-2_test}
    }
    \subfigure['Dale-5', train split]{
        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
        \label{fig:visualizations_dale-5_train}
    }
    \subfigure['Dale-5', test split]{
        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
        \label{fig:visualizations_dale-5_test}
    }
    \caption{Visualization of the models' predictions in the 'Dale' datasets}
    \label{fig:visualizations_dale}
\end{figure}

An interesting pattern appears when doing a qualitative analysis of the models' predictions.
Here, I visualized the predicted coordinates compared to the ground truth coordinates.
Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
As can be seen, the prediction are very precise.
Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are places as green circles.
The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.

The results look very different for the test split.
As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
For all the images, the predictions don't lie on any object.
In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
These findings align with the mean distance scores, described in the sections before.
However, it seems that the model's predictions are all towards the center of the image.
This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
Again, the green circles form the shape of rhombus.
In contrast, the predictions in red almost all cluster in the center of the image.
They form roughly the shape of a smaller rhombus.
This behavior can be observed for all datasets and architectures of the model.
Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
Also the combined visualization shows the same clustering of predictions in the center of the scene, but pattern of the smaller rhombus is better visible.

\begin{figure}[h]
    \centering
    \subfigure['Dale-2', train split]{
        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
        \label{fig:visualizations_dale-2_complete_train}
    }
    \subfigure['Dale-2', test split]{
        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
        \label{fig:visualizations_dale-2_complete_test}
    }
    \subfigure['Dale-5', train split]{
        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
        \label{fig:visualizations_dale-5_complete_train}
    }
    \subfigure['Dale-5', test split]{
        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
        \label{fig:visualizations_dale-5_complete_test}
    }
    \caption{Visualization of the models' predictions in the Dale datasets}
    \label{fig:visualizations_dale_complete}
\end{figure}

These results allow two conclusions.
First, the models are biased to predict coordinates in the center of the image.
The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.

Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
Concluding, the model is able to extract, where objects are located in the image, but can't decide, which of these objects is the target object.

\subsubsection{Bounding box classifier}
\cmtDK{better to call it discriminators?}

One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
Feature extractors like VGG or ResNet are trained on extracting visual information.
In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
To address this challenge, I restructured the task in this experiment.
Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
To do this, fixed-size bounding boxes are extracted around each of the object in the image.
As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
These bounding boxes are shuffled.
Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.

The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
The same is done with the one-hot vector of the target attributes.
In the next step, the dot product between each embedded image and the embedded attributes are calculated.
A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
The loss is calculated using cross entropy.

Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
Again, the model uses a \emph{softmax} function to point to one of these objects.
Since no information about the target object is included, the guesses are expected to be random.

@@

\paragraph*{Results:}
As expected, all baselines achieve a random accuracy for each of the datasets.
The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
This changes however, when information about the target object is included.
Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.

\begin{table}[h]
    \centering
    \begin{tabular}{rcc}
        \toprule
                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
        \textbf{Dale-2}      & 50\%                       & 99\%              \\
        \textbf{Dale-5}      & 22\%                       & 94\%              \\
        \textbf{CLEVR color} & 12\%                       & 93\%              \\
        \bottomrule
    \end{tabular}
    \caption{Results of the bounding box classifier after 10 epochs}
    \label{tab:results_bounding_box_classifier}
\end{table}

The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
This worsens slightly, when more objects are introduced.
More objects lead to more needed attributes to discriminate the objects from each other.
The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
The model can generate these high results, even though its architecture is very simple.
In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.

\subsubsection{Caption generator}

@@

\paragraph*{Results:}
In the next paragraphs, the results of the caption generator will be discussed.
Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
ResNet-3 produced the best results, when it was used as a feature extractor.
The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.

\begin{table}[h]
    \centering
    \begin{tabular}{rccc}
        \toprule
                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
        \bottomrule
    \end{tabular}
    \caption{Results of the caption generator, based on ResNet-3}
    \label{tab:results_caption-generator}
\end{table}

The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
This means that the models produce almost always correct descriptions of an object in the image.
Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
Features of both objects are therefore extracted perfectly and related to the vocabulary.
As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
This changes, when also the masked image is presented to the model.
Then, the accuracy for the target object is with 72\% well above chance.
Still, the wrong descriptions describe almost every time the distractor object.
This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.

When the model is trained on the 'Dale-5' dataset, the results are much worse.
Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
The accuracy stays at the same value.
In contrast, the non-target accuracy is increased by 5\% points.
The model is more likely to identify a wrong object.
Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.

The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.

There are two main possible explanations for the big difference between the two datasets.
First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
The second explanation lies in the used GRE-algorithm.
When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
With four distractors the algorithm is much more likely to produce longer captions.
These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.

\subsection{Language games}
The experiments that are executed using language games have a similar structure as the pre-experiments, since those lied the basis for the language games.

@@

\subsubsection{Discrimination games}
In a discrimination game, the agents are presented with two or more images, one of these being the target image.
The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
The receiver then needs to decide based on the message, which of the images is the target image.

@@

\paragraph*{Results:}
Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
Four different vocabulary sizes $|V|$ are tested.
A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
If the symbols were similarly used, messages would have lengths between one and three symbols.
Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.

The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
Hereby, hidden sizes of 10 and 100 are tested.
On the other hand, three different embedding sizes are tested: 10, 50 and 100.
The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.

\begin{table}[h]
    \centering
    \begin{tabular}{ccccc|cc|cc|cc}
        \toprule
              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {1,99}                              & {24,7\%}                                 & {1}             & {17,3\%}          & {2}             \\
        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {2}                                 & {21,4\%}                                 & {2}             & {17,8\%}          & {1}             \\
        {13}  & {10}    & {10}    & {10}    & {10}    & {94,9\%}                            & {2}                                 & {24,8\%}                                 & {2}             & {17,1\%}          & {2}             \\
        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {2}                                 & {21,9\%}                                 & {2}             & {17,9\%}          & {1}             \\
        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {1}                                 & {23\%}                                   & {2}             & {15,9\%}          & {2}             \\
        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {2}                                 & {24\%}                                   & {2}             & {18,1\%}          & {2}             \\
        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {2}                                 & {24,4\%}                                 & {2}             & {15,8\%}          & {2}             \\
        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {1}                                 & {25,3\%}                                 & {2}             & {15,6\%}          & {1}             \\
        \bottomrule
    \end{tabular}
    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
    \label{tab:results_discriminator}
\end{table}

For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
The receiver identifies almost every sample correctly when all sizes are 10.
When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.

The results change, when using the \emph{Dale-5} dataset with four distractors.
With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.

\cmtDK{calculation of loss?}
\cmtDK{length}

Two conclusions can be drawn.
First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g. from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.

The second conclusion that can be drawn looks at the differences between the two datasets.
Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
There are 48 possible combinations of attributes.
Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.

\cmtDK{look at language}
\cmtDK{similarity to bounding box classifier}

\subsubsection{Caption generator games}
In a next step, it is tested if the agents can learn to extract features of the objects together.
For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
The receiver only sees the image without additional information.
\cmtDK{human bias vs emergent language}

@@

\paragraph*{Results:}
In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
In every configuration of the variables, the agents achieve a very high performance.
The worst average distance across the test dataset is 10 pixels, which still points on an object.
Also the accuracy, which evaluates how many guesses of the receiver were pointing on an object reflects this fact.
All configuration achieve an accuracy higher than 96,7\%.
This aligns also with the results from the single neural models, where the average distance was similarly low.
In general, this shows that the agents are able to predict coordinates together.
However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
When a different sender is used that is receiving the masked image as input, the results indicate that the message has in fact an influence and the receiver learns how to interpret the message.
The results for this setup are summarized in \ref{tab:results_masked_predictor_game}.
In three of six experiments, the agents predict much higher average distances.
In one case, the mean distance lies at 19,8 pixels, while the accuracy is only at 55,1\%.
Since the receiver is the same in both setups, this observation indicates that a meaningful message is communicated between the agents.
